---
title: "Financial Options: A Monte Carlo Approach"
author: "Group 6: Trevor Isaacson, Noah Smith and Buzz Walter "
date: "12/7/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
set.seed(400)
library(tidyr)
library(tidyverse)
library(rmarkdown)
library(ggplot2)
library(dbplyr)
library(knitr)
```

## Background

- Our project is based on a paper called "Options: A Monte Carlo Approach"
- Written by Phelim P. Boyle in 1977 and published in the Journal of Financial Economics
- Boyle is an Irish economist and a pioneer of quantitative finance
- He is best known for his use of Monte Carlo methods in option pricing

## Financial Terminology

- Stock: security ownership in a particular company 
  - ex: common stock of Apple is ~ $165
- Dividend: A sum of money paid regularly by a company to its shareholders
- Inflation Rate: rate at which the value of a currency is falling and the general level of prices for goods and services is rising
- Treasury Yield: yield of a US Treasury Bond matching investment duration
- Risk Free Rate: Represents the interest an investor would expect from a risk-free investment over a specific time
  - RFr = Inflation Rate - Treasury Yield
  
## The LogNormal Distribution

- Distribution of logarithmic values from a related normal distribution
- Allow for only positive variables whereas the normal distribution can allow for negative variables
- More useful in finance because the log normal curve can better identify the compounding return a stock can achieve over a long period of time

## The LogNormal Distribution

```{r warning=FALSE, echo = FALSE}
sample1 <- rnorm(n=1e4, mean=1, sd=1)
sample2 <- rlnorm(n=1e4, mean=1, sd=1)
lognorms <- data.frame(Value = c(sample1, sample2), x_bar = factor(c(rep("Normal", length(sample1)), rep("LogNormal", length(sample2)))))
ggplot(data = lognorms, aes(x = Value, group = x_bar, fill = x_bar, color = x_bar)) +
  geom_density(adjust = 1.5, alpha = .4) + scale_x_continuous(limits =c(0, 10)) + ggtitle("LogNormal Distributions")
```


## What are options?

- Options are a very powerful and influential tool used heavily in financial markets
- It is a contract that is valued based on an underlying asset or security like a common stock or bond
  - Derivative, price is linked to price of underlying asset
- Options are more risky than common stock as the price moves faster
- Investors use options for income, to hedge risk and to speculate about future market movements
- When investors sell an underlying stock as an option they receive a premium up front for selling the contract aka the option 

## How do options work?

- A single options contract represents 100 shares of the underlying asset it is representing
- Contract gives the buyer the right to buy the underlying asset at a specific price on or by a certain date
- Valuing options is essentially about determining the probabilities of future prices
  - The more likely something is to occur, the more expensive an option for that event
  - The less time to maturity, the less valuable the option 
  - As time to maturity decreases on an option the value of the option will decrease as the delta variable for time decreases assuming value of the stock stays constant.
  - Previous to these methods, there was no agreed upon method to determine option prices 
  
## European vs American Options

- There are two types of financial options: European and American
- The main thing to remember between the two, American options can be exercised at any time
- European options can only be exercised at the time of expiration
- Again a Contract gives the buyer the right to buy the underlying asset at a specific price on or by a certain date and exercising the contract just means you finish the translation and buy the stocks.


  
## Purpose of the Paper

- To show that Monte Carlo simulations provide a method to obtaining numerical solutions to option valuation problems
- Option valuation models are important because large amounts of corporate liabilities can be expressed in terms of options
- Two other methods exist and this method should be used as a supplement
- Boyle's method is simple and flexible as it can be easily modified for different processes


## Black-Scholes Model

- The Black-Scholes model is the basis for the model from this paper
- Developed in 1973 it is one of the best methods for pricing an option contract
- The paper that we replicated derived their methods from the Black-Scholes Model

C = $S_t*N(d_1)-K_e^{-rt}*N(d_2)$

$d_1 = \frac{ln({S_t/K})+(r+\sigma^2/2)*t}{\sigma_s\sqrt{t}}$

$d_2=d_1-\sigma_s*\sqrt{t}$

## Black-Scholes Model Cont.

- C = call option price
- S = Current underling stock price
- K = Strike price
- r = Risk-free interest rate
- t = Time to maturity
- N = A normal distribution

- This model only prices European options since again American options can be exercised before expiration date causing for more variability.
- This model assumes that variability remains constant.


## Method Overview

- Our Monte Carlo method is used to obtain estimates of option values on dividend paying stocks for European Call options
- Method uses the fact that the distribution of terminal stock prices is determined by the process generating future stock price movements
- This process can then be simulated to generate a series of stock price trajectories
- The series determines a set of terminal stock values which can be used to obtain an estimate of the option value


## Monte Carlo

- $\int f(y)g(y)$
- $\hat{g}$ = MAX[$S_{T+1}$ - E, 0]
- $f(y)$ = lognormal distribution
- $g(y)$ = result we get for final option price
  - random variable because we include $\tilde{x}$


## How it works

- $S_t$ = current stock price at time t    
- r = risk free rate per quarter compounded continuously    
- $\sigma^2$ = the (assumed constant) variance rate per quarter on the underlying stock    
- $D_t$ = the dividend payable at time t   
- E = exercise price of the option    

- Using the properties of the lognormal distribution, we can generate the distribution of stock prices one period hence by forming the random variable, 
  - $S_{t+1} = S_t*exp[r - \sigma^2/2 + \sigma*\tilde{x}]$,     $\tilde{x}$ is $N(0, 1)$


## Simulation

- To set up the simulation method in this case, a value of $S_{t+1}$ is generated
- If $S_{t+1}$ is greater than $D_{t+1}$, then ($S_{t+1}$ - $D_{t+1}$) is used as the initial value at the start of the second period  
  - The procedure is continued until a value of $S_t$ is obtained   
- If at some stage $S_{t+m} (m = 1,2,... (T-t-1)$ is less than or equal to the corresponding dividend payment $D_{t+m}$ the process stops    
  - In this case, another simulation trial is started again from time t   
  
  
## Simulation 

- A series of simulation trials is carried out in this way and the expected value of $MAX[S_t-E, 0]$ is found   
  - This quantity is then discounted as the risk-free rate to yield an estimate of the option value $\hat{g}$

## Simulation Code    

- Trials = 5000
- Periods to Maturity = c(2,4,6,8,10,12,14,16,18,20)
- stock_prices = c(25, 50, 75)
- Exercise Price = 50
- Risk Free Rate = 0.015
- $\sigma^2$ = 0.025 
- Dividend = 0.25


```{r echo = FALSE}
trials = 5000
periods_to_maturity = c(2,4,6,8,10,12,14,16,18,20)
stock_prices = c(25, 50, 75)
stockExerciseRatio = c(0.5, 1, 1.25)
E = 50
r = 0.015
sigma_squared = 0.025 
dividend = 0.25

#chosen from October 2021, but others can be chosen to try to match historical data
yield_points = c(.09,.1,.11,.16,.36)
inflation_rate = 0.062

#rate to be used to evaluate the present day value of an investment
risk_free_rate = function(inflation_rate, period_to_maturity){
  quarters = c(2,4,8,12,20)
  interpolate_yield_curve = approxfun(x = quarters, y = yield_points)
  return(inflation_rate - interpolate_yield_curve(period_to_maturity))
}


# an individual evaluation of a trial for a single option valuation problem
generate_option_val = function(stock_price, r, period_to_maturity, sigma){
  x_tilda = rnorm(periods_to_maturity, mean = 0, sd = 1)
  for (i in 1:period_to_maturity){
    new_q_price = stock_price * exp((r - (sigma^2)/2) + sigma * x_tilda[i])
    #case where trial needs to be restarted from price at quarter t rather than quarter 1(see p. 329 of Boyle(1977))
    if (new_q_price -  dividend <= 0){
      stock_price = generate_option_val(stock_price, r, period_to_maturity - i + 1, sigma)
      break 
    }
    #case where we need to pay out dividends to evaluate new stock price
    else{
      stock_price = new_q_price - dividend
    }
  }
  future_option_value = max(c(stock_price - E, 0))
  return(future_option_value - risk_free_rate(inflation_rate,period_to_maturity)*future_option_value)
}

#simulation looking at the case of 2 quarters to maturity and a stock price of 25 matching
option_vals = rep(NA, trials)
for (i in 1:trials){
  option_vals[i] = generate_option_val(stock_prices[1], r, periods_to_maturity[1],sqrt(sigma_squared))
}
# print(paste("option value: ",mean(option_vals)))
# print(paste("se of estimate: ",sd(option_vals)/(sqrt(trials))))
```

```{r echo = FALSE, tidy = TRUE}
set.seed(400)
trials = 5000
periods_to_maturity = c(2,4,6,8,10,12,14,16,18,20)
stock_prices = c(25, 50, 75)
stockExerciseRatio = c(0.5, 1, 1.5)
E = 50
r = 0.015
sigma = 0.025 
dividend = 0.25

#chosen from October 2021, but others can be chosen to try to match historical data
yield_points = c(.09,.1,.11,.16,.36)
#yield_points = c(.04,.06,.9,.10,.14)
#inflation_rate = .1585
inflation_rate = 0.062

#rate to be used to evaluate the present day value of an investment
risk_free_rate = function(inflation_rate, periods_to_maturity){
  quarters = c(2,4,8,12,20)
  interpolate_yield_curve = approxfun(x = quarters, y = yield_points)
  return(inflation_rate - interpolate_yield_curve(periods_to_maturity))
}


# an individual evaluation of a trial for a single option valuation problem
generate_option_val = function(stock_price, r, periods_to_maturity, sigma){
  x_tilda = rnorm(periods_to_maturity, mean = 0, sd = 1)
  for (i in 1:periods_to_maturity){
    new_q_price = stock_price * exp((r - (sigma^2)/2) + sigma * x_tilda[i])
    #case where trial needs to be restarted from price at quarter t rather than quarter 1(see p. 329 of Boyle(1977))
    if (new_q_price -  dividend <= 0){
      stock_price = generate_option_val(stock_price, r, periods_to_maturity - i + 1, sigma)
      break 
    }
    #case where we need to pay out dividends to evaluate new stock price
    else{
      stock_price = new_q_price - dividend
    }
  }
  future_option_value = max(c(stock_price - E, 0))
  return(future_option_value - risk_free_rate(inflation_rate,periods_to_maturity)*future_option_value)
}

#simulation looking at the case of 2 quarters to maturity and a stock price of 25 matching
option_vals = rep(NA, trials)
newmean = c()
newsd = c()
option_mean_25 = c()
option_sd_25 = c()

## need this to make variable option_vals again but empty so numbers dont add to each other
option_vals = rep(NA, trials)

for (l in 1:9){
  for (i in 1:trials){
    option_vals[i] = generate_option_val(stock_prices[1], r, periods_to_maturity[l],sqrt(sigma))
  }
  option_mean_25 = append(option_mean_25,mean(option_vals))
  newsd = sd(option_vals)/(sqrt(trials))
  option_sd_25 = append(option_sd_25,newsd)
}

## now this is for stock prices at 50

option_vals = rep(NA, trials)
newmean = c()
newsd = c()
option_mean_50 = c()
option_sd_50 = c()
for (l in 1:9){
  for (i in 1:trials){
    option_vals[i] = generate_option_val(stock_prices[2], r, periods_to_maturity[l],sqrt(sigma))
  }
  option_mean_50 = append(option_mean_50,mean(option_vals))
  newsd = sd(option_vals)/(sqrt(trials))
  option_sd_50 = append(option_sd_50,newsd)
}


## now this is for stock prices at 75
option_vals = rep(NA, trials)
newmean = c()
newsd = c()
option_mean_75 = c()
option_sd_75 = c()
for (l in 1:9){
  for (i in 1:trials){
    option_vals[i] = generate_option_val(stock_prices[3], r, periods_to_maturity[l], sqrt(sigma))
  }
  option_mean_75 = append(option_mean_75,mean(option_vals))
  newsd = sd(option_vals)/(sqrt(trials))
  option_sd_75 = append(option_sd_75,newsd)
}

periods_to_maturity =c(2,4,6,8,10,12,14,16,18)

df_25 <- data.frame(Periods_til_maturity = periods_to_maturity, Option_values=option_mean_25, Standard_deviation=option_sd_25, lower_95_CI = option_mean_25 - 2*option_sd_25, upper_95_CI = option_mean_25 + 2*option_sd_25)
df_25 = round(df_25, 4)

df_50 <- data.frame(Periods_til_maturity = periods_to_maturity, Option_values=option_mean_50, Standard_deviation=option_sd_50, lower_95_CI = option_mean_50 - 2*option_sd_50, upper_95_CI = option_mean_50 + 2*option_sd_50)
df_50 = round(df_50, 4)

df_75 <- data.frame(Periods_til_maturity = periods_to_maturity, Option_values=option_mean_75,Standard_deviation=option_sd_75, lower_95_CI = option_mean_75 - 2*option_sd_75, upper_95_CI = option_mean_75 + 2*option_sd_75)
df_75 = round(df_75, 4)

```

## Stock Price = $25

```{r fig.height= 5}
df_25 %>% kable()

```

## Stock Price = $25

```{r warning = FALSE}
real_values_25 = c(0.004, 0.087, 0.273, 0.539, 0.879, 1.202, 1.547, 1.897, 2.281)
table1_25 = data.frame(x = periods_to_maturity, y = real_values_25)

real_values_50 = c(5.121, 7.427, 9.247, 10.793, 11.923, 13.135, 14.281, 15.079, 15.945)
table1_50 = data.frame(x = periods_to_maturity, y = real_values_50)

real_values_75 = c(26.445, 28.250, 29.832, 31.257, 32.319, 33.491, 34.835, 35.503, 36.382)
table1_75 = data.frame(x = periods_to_maturity, y = real_values_75)

ggplot() + 
  geom_point(aes(x=Periods_til_maturity, y=Option_values), df_25) + 
  geom_line(aes(x=Periods_til_maturity, y=Option_values), df_25) +
  geom_ribbon(aes(x=Periods_til_maturity, ymin = lower_95_CI, ymax = upper_95_CI), df_25, alpha = 0.3) + 
  geom_point(aes(x = x, y = y), data = table1_25, color = "red", alpha = 0.5) +
  geom_line(aes(x = x, y = y), data = table1_25, color = "red", alpha = 0.5) +
  xlab("Number of Periods to Maturity")+ ylab("Option values by Crude Monte Carlo")+
  ggtitle("Options at Stock Price of 25")

```

## Stock Price = $50

```{r fig.height= 5}
df_50 %>% kable()

```


## Stock Price = $50

```{r}
ggplot(df_50, aes(x=Periods_til_maturity, y=Option_values)) + 
  geom_point() + geom_line()+
  geom_ribbon(aes(ymin = lower_95_CI, ymax = upper_95_CI), alpha = 0.3) + 
  geom_point(aes(x = x, y = y), data = table1_50, color = "red", alpha = 0.5) +
  geom_line(aes(x = x, y = y), data = table1_50, color = "red", alpha = 0.5) +
  xlab("Number of Periods to Maturity")+
  ylab("Option values by Crude Monte Carlo")+
  ggtitle("Options at Stock Price of 50")
```

## Stock Price = $75



```{r fig.height= 5}
df_75 %>% kable()

```


## Stock Price = $75


```{r}
ggplot(df_75, aes(x=Periods_til_maturity, y=Option_values)) + 
  geom_point() + geom_line() +
  geom_ribbon(aes(ymin = lower_95_CI, ymax = upper_95_CI), alpha = 0.3) + 
  geom_point(aes(x = x, y = y), data = table1_75, color = "red", alpha = 0.5) +
  geom_line(aes(x = x, y = y), data = table1_75, color = "red", alpha = 0.5) +
  xlab("Number of Periods to Maturity")+ ylab("Option values by Crude Monte Carlo") +
  ggtitle("Options at Stock Price of 75")

```


## Questions?




## References
https://www.investopedia.com/options-basics-tutorial-4583012

https://www.investopedia.com/terms/r/risk-freerate.asp

https://www.investopedia.com/terms/b/blackscholes.asp

https://www.in2013dollars.com/UK-inflation-rate-in-1977
