---
title: "Financial Options: A Monte Carlo Approach"
author: "Group 6: Trevor Isaacson, Noah Smith and Buzz Walter "
date: "12/7/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
set.seed(400)
library(tidyr)
library(tidyverse)
library(rmarkdown)
library(ggplot2)
library(dbplyr)
```

## Background

- Our project is based on a paper called "Options: A Monte Carlo Approach"
- Written by Phelim P. Boyle in 1977 and published in the Journal of Financial Economics
- Boyle is an Irish economist and a pioneer of quantitative finance
- He is best known for his use of Monte Carlo methods in option pricing

## Financial Terminology

- Stock: security ownership in a particular company 
  - ex: common stock of Apple is ~ $165
- Inflation Rate: rate at which the value of a currency is falling and the general level of prices for goods and services is rising
- Treasury Yield: yield of a US Treasury Bond matching investment duration
- Risk Free Rate: Represents the interest an investor would expect from a risk-free investment over a specific time
  - RFr = Treasury Yield - Inflation Rate
  
## The LogNormal Distribution




## The LogNormal Distribution

```{r warning=FALSE, echo = FALSE}
sample1 <- rlnorm(n=1e4, mean=0, sd=1)
sample2 <- rlnorm(n=1e4, mean=1, sd=1)
sample3 <- rlnorm(n=1e4, mean=2, sd=1)
lognorms <- data.frame(Value = c(sample1, sample2, sample3), x_bar = factor(c(rep(0, length(sample1)), rep(1, length(sample2)), rep(2, length(sample3)))))
ggplot(data = lognorms, aes(x = Value, group = x_bar, fill = x_bar, color = x_bar)) +
  geom_density(adjust = 1.5, alpha = .4) + scale_x_continuous(limits =c(0, 25)) + ggtitle("LogNormal Distributions")
```


## What are options?

- Options are a very powerful and influential tool used heavily in financial markets
- It is a contract that is valued based on an underlying asset or security like a common stock or bond
  - Derivative, price is linked to price of underlying asset
- Options are more risky than common stock as the price moves faster
- Investors use options for income, to hedge risk and to speculate about future market movements

## How do options work?

- A single options contract represents 100 shares of the underlying asset it is representing
- Contract gives the buyer the right to buy the underlying asset at a specific price on a certain date
- Valuing options is essentially about determining the probabilities of future prices
  - The more likely something is to occur, the more expensive an option for that event
  - The less time to maturity, the less valuable the option 
  - Previous to these methods, there was no agreed upon method to determine option prices 
  
## European vs American Options

There are two types of financial options: European and American
- 

  
## Purpose of the Paper

- To show that Monte Carlo simulations provide a method to obtaining numerical solutions to option valuation problems
- Option valuation models are important because large amounts of corporate liabilities can be expressed in terms of options
- Two other methods exist and this method should be used as a supplement
- Boyle's method is simple and flexible as it can be easily modified for different processes


## Monte Carlo

- $\hat{g}$ = MAX[$S_{t+1}$ - E, 0]
- $f(y)$ = lognormal distribution
- $g(y)$ = result we get for final stock price
  - random variable because we include x tilda


## Black Scholes Model





## Method Overview

- Our Monte Carlo method is used to obtain estimates of option values on dividend paying stocks for European Call options
- Method uses the fact that the distribution of terminal stock prices is determined by the process generating future stock price movements
- This process can then be simulated to generate a series of stock price trajectories
- The series determines a set of terminal stock values which can be used to obtain an estimate of the option value

## How it works




## Simulation
```{r echo = FALSE}
trials = 5000
periods_to_maturity = c(2,4,6,8,10,12,14,16,18,20)
stock_prices = c(25, 50, 75)
stockExerciseRatio = c(0.5, 1, 1.5)
E = 50
r = 0.015
sigma = 0.025 
dividend = 0.25
#chosen from October 2021, but others can be chosen to try to match historical data
yield_points = c(.09,.1,.11,.16,.36)
#yield_points = c(.04,.06,.9,.10,.14)
#inflation_rate = .1585
inflation_rate = 0.062

#rate to be used to evaluate the present day value of an investment
risk_free_rate = function(inflation_rate, periods_to_maturity){
  quarters = c(2,4,8,12,20)
  interpolate_yield_curve = approxfun(x = quarters, y = yield_points)
  return(inflation_rate - interpolate_yield_curve(periods_to_maturity))
}


# an individual evaluation of a trial for a single option valuation problem
generate_option_val = function(stock_price, r, periods_to_maturity, sigma){
  x_tilda = rnorm(periods_to_maturity, mean = 0, sd = 1)
  for (i in 1:periods_to_maturity){
    new_q_price = stock_price * exp((r - (sigma^2)/2) + sigma * x_tilda[i])
    #case where trial needs to be restarted from price at quarter t rather than quarter 1(see p. 329 of Boyle(1977))
    if (new_q_price -  dividend <= 0){
      stock_price = generate_option_val(stock_price, r, periods_to_maturity - i + 1, sigma)
      break 
    }
    #case where we need to pay out dividends to evaluate new stock price
    else{
      stock_price = new_q_price - dividend
    }
  }
  future_option_value = max(c(stock_price - E, 0))
  return(future_option_value - risk_free_rate(inflation_rate,periods_to_maturity)*future_option_value)
}

#simulation looking at the case of 2 quarters to maturity and a stock price of 25 matching
option_vals = rep(NA, trials)
newmean = c()
newsd = c()
option_mean_25 = c()
option_sd_25 = c()

## need this to make variable option_vals again but empty so numbers dont add to each other
option_vals = rep(NA, trials)

for (l in 1:9){
  for (i in 1:trials){
    option_vals[i] = generate_option_val(stock_prices[1], r, periods_to_maturity[l], sqrt(sigma))
  }
  print(periods_to_maturity[l])
  option_mean_25 = append(option_mean_25, mean(option_vals))
  newsd = sd(option_vals)/(sqrt(trials))
  option_sd_25 = append(option_sd_25,newsd)
}

## now this is for stock prices at 50

option_vals = rep(NA, trials)
newmean = c()
newsd = c()
option_mean_50 = c()
option_sd_50 = c()
for (l in 1:9){
  for (i in 1:trials){
    option_vals[i] = generate_option_val(stock_prices[2], r, periods_to_maturity[l],sqrt(sigma))
  }
  option_mean_50 = append(option_mean_50,mean(option_vals))
  newsd = sd(option_vals)/(sqrt(trials))
  option_sd_50 = append(option_sd_50,newsd)
}


## now this is for stock prices at 75
option_vals = rep(NA, trials)
newmean = c()
newsd = c()
option_mean_75 = c()
option_sd_75 = c()
for (l in 1:9){
  for (i in 1:trials){
    option_vals[i] = generate_option_val(stock_prices[3], r, periods_to_maturity[l],sqrt(sigma))
  }
  option_mean_75 = append(option_mean_75,mean(option_vals))
  newsd = sd(option_vals)/(sqrt(trials))
  option_sd_75 = append(option_sd_75,newsd)
}



df_25 <- data.frame(Periods_til_maturity = periods_to_maturity, Option_values=option_mean_25, Standard_deviation=option_sd_25, lower_95_CI = option_mean_25 - 2*option_sd_25, upper_95_CI = option_mean_25 + 2*option_sd_25)
df_25 = round(df_25, 4)
df_25

df_50 <- data.frame(Periods_til_maturity = periods_to_maturity, Option_values=option_mean_50, Standard_deviation=option_sd_50, lower_95_CI = option_mean_50 - 2*option_sd_50, upper_95_CI = option_mean_50 + 2*option_sd_50)
df_50 = round(df_50, 4)
df_50

df_75 <- data.frame(Periods_til_maturity = periods_to_maturity, Option_values=option_mean_75,Standard_deviation=option_sd_75, lower_95_CI = option_mean_75 - 2*option_sd_75, upper_95_CI = option_mean_75 + 2*option_sd_75)
df_75 = round(df_75, 4)
df_75
```














## References
https://www.investopedia.com/options-basics-tutorial-4583012