---
title: "Financial Options: A Monte Carlo Approach"
author: "Group 6: Trevor Isaacson, Noah Smith and Buzz Walter "
date: "12/7/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
set.seed(400)
library(tidyr)
library(tidyverse)
library(rmarkdown)
library(ggplot2)
library(dbplyr)
```

## Background

- Our project is based on a paper called "Options: A Monte Carlo Approach"
- Written by Phelim P. Boyle in 1977 and published in the Journal of Financial Economics
- Boyle is an Irish economist and a pioneer of quantitative finance
- He is best known for his use of Monte Carlo methods in option pricing

## What are options?

- Options are a very powerful and influential tool used heavily in financial markets
- It is a contract that is valued based on an underlying asset or security like a common stock or bond
  - Derivative, price is linked to price of underlying asset
- Options are more risky than common stock as the price moves faster
- Investors use options for income, to hedge risk and to speculate about future market movements

## How do options work?

- A single options contract represents 100 shares of the underlying asset it is representing
- Contract gives the buyer the right to buy the underlying asset at a specific price on a certain date
- 




## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```










## References
https://www.investopedia.com/options-basics-tutorial-4583012