---
title: "Financial Options: A Monte Carlo Approach"
author: "Group 6: Trevor Isaacson, Noah Smith and Buzz Walter "
date: "12/7/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
set.seed(400)
library(tidyr)
library(tidyverse)
library(rmarkdown)
library(ggplot2)
library(dbplyr)
```

## Background

- Our project is based on a paper called "Options: A Monte Carlo Approach"
- Written by Phelim P. Boyle in 1977 and published in the Journal of Financial Economics
- Boyle is an Irish economist and a pioneer of quantitative finance
- He is best known for his use of Monte Carlo methods in option pricing

## What are options?

- Options are a very powerful and influential tool used heavily in financial markets
- It is a contract that is valued based on an underlying asset or security like a common stock or bond
  - Derivative, price is linked to price of underlying asset
- Options are more risky than common stock as the price moves faster
- Investors use options for income, to hedge risk and to speculate about future market movements
- When investors sell an underlying stock as an option they receive a premium up front for selling the contract aka the option 

## How do options work?

- A single options contract represents 100 shares of the underlying asset it is representing
- Contract gives the buyer the right to buy the underlying asset at a specific price on or by a certain date
- Valuing options is essentially about determining the probabilities of future prices
  - The more likely something is to occur, the more expensive an option for that event
  - The less time to maturity, the less valuable the option 
  - As time to maturity decreases on an option the value of the option will decrease as the delta variable for time decreases assuming value of the stock stays constant.
  - 
  
## Purpose of the Paper

- To show that Monte Carlo simulations provide a method to obtaining numerical solutions to option valuation problems
- Two other methods exist and this method should be used as a supplement
- Boyle's method is simple and flexible as it can be easily modified for different processes

## Overview

- Method uses the fact that the distribution of terminal stock prices is determined by the process generating future stock price movements
- 


## Slide with R Output

```{r functions, include=FALSE}
#Back group code this should be include= False

set.seed(400)
trials = 5000
periods_to_maturity = c(2,4,6,8,10,12,14,16,18,20)
stock_prices = c(25, 50, 75)
stockExerciseRatio = c(0.5, 1, 1.5)
E = 50
r = 0.015
sigma = 0.025 
dividend = 0.25

#chosen from October 2021, but others can be chosen to try to match historical data
yield_points = c(.09,.1,.11,.16,.36)
#yield_points = c(.04,.06,.9,.10,.14)
#inflation_rate = .1585
inflation_rate = 0.062

#rate to be used to evaluate the present day value of an investment
risk_free_rate = function(inflation_rate, periods_to_maturity){
  quarters = c(2,4,8,12,20)
  interpolate_yield_curve = approxfun(x = quarters, y = yield_points)
  return(inflation_rate - interpolate_yield_curve(periods_to_maturity))
}


# an individual evaluation of a trial for a single option valuation problem
generate_option_val = function(stock_price, r, periods_to_maturity, sigma){
  x_tilda = rnorm(periods_to_maturity, mean = 0, sd = 1)
  for (i in 1:periods_to_maturity){
    new_q_price = stock_price * exp((r - (sigma^2)/2) + sigma * x_tilda[i])
    #case where trial needs to be restarted from price at quarter t rather than quarter 1(see p. 329 of Boyle(1977))
    if (new_q_price -  dividend <= 0){
      stock_price = generate_stock_price(stock_price, r, periods_to_maturity - i + 1, sigma)
      break 
    }
    #case where we need to pay out dividends to evaluate new stock price
    else{
      stock_price = new_q_price - dividend
    }
  }
  future_option_value = max(c(stock_price - E, 0))
  return(future_option_value - risk_free_rate(inflation_rate,periods_to_maturity)*future_option_value)
}

#simulation looking at the case of 2 quarters to maturity and a stock price of 25 matching
option_vals = rep(NA, trials)
newmean = c()
newsd = c()
option_mean_25 = c()
option_sd_25 = c()

## need this to make variable option_vals again but empty so numbers dont add to each other
option_vals = rep(NA, trials)

for (l in 1:9){
  for (i in 1:trials){
  option_vals[i] = generate_option_val(stock_prices[1], r, periods_to_maturity[l],sqrt(sigma))


}
  option_mean_25 = append(option_mean_25,mean(option_vals))
  newsd = sd(option_vals)/(sqrt(trials))
  option_sd_25 = append(option_sd_25,newsd)
}



## now this is for stock prices at 50

option_vals = rep(NA, trials)
newmean = c()
newsd = c()
option_mean_50 = c()
option_sd_50 = c()
for (l in 1:9){
  for (i in 1:trials){
  option_vals[i] = generate_option_val(stock_prices[2], r, periods_to_maturity[l],sqrt(sigma))


}
  option_mean_50 = append(option_mean_50,mean(option_vals))
  newsd = sd(option_vals)/(sqrt(trials))
  option_sd_50 = append(option_sd_50,newsd)
}




## now this is for stock prices at 75

option_vals = rep(NA, trials)
newmean = c()
newsd = c()
option_mean_75 = c()
option_sd_75 = c()
for (l in 1:9){
  for (i in 1:trials){
  option_vals[i] = generate_option_val(stock_prices[3], r, periods_to_maturity[l],sqrt(sigma))


}
  option_mean_75 = append(option_mean_75,mean(option_vals))
  newsd = sd(option_vals)/(sqrt(trials))
  option_sd_75 = append(option_sd_75,newsd)
}





df_25 <- data.frame(Periods_till_maturity=c(2,4,6,8,10,12,14,16,18),
  Option_values=option_mean_25,
                 Standard_deviation=option_sd_25)

df_25
df_50 <- data.frame(Periods_till_maturity=c(2,4,6,8,10,12,14,16,18),
  Option_values=option_mean_50,
                 Standard_deviation=option_sd_50)

df_50
df_75 <- data.frame(Periods_till_maturity=c(2,4,6,8,10,12,14,16,18),
  Option_values=option_mean_75,
                 Standard_deviation=option_sd_75)

df_75


```

## Slide with Plot

```{r pressure}
df_25
df_50
df_75
```










## References
https://www.investopedia.com/options-basics-tutorial-4583012

https://www.investopedia.com/terms/r/risk-freerate.asp

https://www.investopedia.com/terms/b/blackscholes.asp

https://www.in2013dollars.com/UK-inflation-rate-in-1977
