---
title: "project research"
author: "Trevor Isaacson, Noah Smith, Buzz Walter"
date: "12/7/2021"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyr)
library(tidyverse)
library(rmarkdown)
library(ggplot2)
library(dbplyr)
```

## This document is for any research or progress made on the project and is not meant to be the final result


St = current stock price at time t
r = risk free rate per quarter compounded continuously
sigma_squared = the (assumed constant) variance rate per quarter on the underlying stock
Dt = the dividend payable at time t
E = exercise price of the option
T = exercise date of the option

Expected Return on the Stock = $E(S_T / S_t) = exp(r[T - time])$
  Assumed that the ratio of successive stock values follows a lognormal distribution
Using the properties of the lognormal distribution, this means that we can generate the distribution of stock prices one period hence by forming the RV, 
  $S_t+1 = S_t*exp[r - sigma^2/2 + sigma*\tilde{x}] $ 
  where $\tilde{x}$ is a normally distributed random variable with mean = 0 and unit variance
  It is assumed that time is measured in units of one quarter and $S_t$ represents the stock price just after the quarterly dividend $D_t$ has been paid
  
To set up the simulation method in this case, a value of $S_t+1$ is generated.  
If this value is greater than $D_t+1$, then ($S_t+1$ - $D_t+1$) is used as the initial value at the start of the second period
  The procedure is continued until a value of $S_r$ is obtained
If at some stage $S_t+m (m = 1,2,... (T-time-1)$ is less than or equal to the corresponding dividend payment $D_t+m$ the process stops
  In this case, another simulation trial is started again from time t
A series of simulation trials is carried out in this way and the expected value of $MAX[S_r-E, 0]$ is found
  This quantity is then discounted as the risk-free rate to yield an estimate of the option value $\hat{g}$
  95% CIs are given by $\hat{g}$ +- $2*\hat{s}/ sqrt(n)$
  
Recreation of Monte Carlo Simulation of Table 1 in Paper:
```{r}
trials = 5000
periods_to_maturity = c(2,4,6,8,10,12,14,16,18,20)
stock_prices = c(25, 50, 75)





```

