---
title: "project research"
author: "Trevor Isaacson, Noah Smith, Buzz Walter"
date: "12/7/2021"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
set.seed(400)
library(tidyr)
library(tidyverse)
library(rmarkdown)
library(ggplot2)
library(dbplyr)
```

$S_t$ = current stock price at time t    
r = risk free rate per quarter compounded continuously    
$\sigma^2$ = the (assumed constant) variance rate per quarter on the underlying stock    
$D_t$ = the dividend payable at time t   
E = exercise price of the option    
T = exercise date of the option   

Expected Return on the Stock = $E(S_T / S_t) = exp(r[T - time])$    

* Assumed that the ratio of successive stock values follows a lognormal distribution due to the assumed normal distribution of rates of return. Black-Scholes model assumes this distribution. 
* Using the properties of the lognormal distribution, this means that we can generate the distribution of stock prices one period hence by forming the random variable, 
  $S_{t+1} = S_t*exp[r - \sigma^2/2 + \sigma*\tilde{x}]$
  - Where $\tilde{x}$ is a normally distributed random variable with mean = 0 and unit variance         
* It is assumed that time is measured in units of one quarter and $S_t$ represents the stock price just after the quarterly dividend $D_t$ has been paid        
  
To set up the simulation method in this case, a value of $S_{t+1}$ is generated.    

* If this value is greater than $D_{t+1}$, then ($S_{t+1}$ - $D_{t+1}$) is used as the initial value at the start of the second period    
  - The procedure is continued until a value of $S_t$ is obtained   
* If at some stage $S_{t+m} (m = 1,2,... (T-t-1)$ is less than or equal to the corresponding dividend payment $D_{t+m}$ the process stops    
  - In this case, another simulation trial is started again from time t   
A series of simulation trials is carried out in this way and the expected value of $MAX[S_r-E, 0]$ is found   
  - This quantity is then discounted as the risk-free rate to yield an estimate of the option value $\hat{g}$     
  - 95% CIs are given by $\hat{g}$ +- $2*\hat{s}/ sqrt(n)$  
  
**Recreation of Monte Carlo Simulation of Table 1 in Paper:**   

Stock prices of 25, 50 and 75 lead to Stock/Exercise ratios = 0.5, 1, and 1.5
```{r}

trials = 5000
periods_to_maturity = c(2,4,6,8,10,12,14,16,18,20)
stock_prices = c(25, 50, 75)
stockExerciseRatio = c(0.5, 1, 1.25)
E = 50
r = 0.015
sigma_squared = 0.025 
dividend = 0.25

#chosen from October 2021, but others can be chosen to try to match historical data
yield_points = c(.09,.1,.11,.16,.36)

inflation_rate = 0.062

#rate to be used to evaluate the present day value of an investment
risk_free_rate = function(inflation_rate, period_to_maturity){
  quarters = c(2,4,8,12,20)
  interpolate_yield_curve = approxfun(x = quarters, y = yield_points)
  return(inflation_rate - interpolate_yield_curve(period_to_maturity))
}


# an individual evaluation of a trial for a single option valuation problem
generate_option_val = function(stock_price, r, period_to_maturity, sigma){
  x_tilda = rnorm(periods_to_maturity, mean = 0, sd = 1)
  for (i in 1:period_to_maturity){
    new_q_price = stock_price * exp((r - (sigma^2)/2) + sigma * x_tilda[i])
    #case where trial needs to be restarted from price at quarter t rather than quarter 1(see p. 329 of Boyle(1977))
    if (new_q_price -  dividend <= 0){
      stock_price = generate_stock_price(stock_price, r, period_to_maturity - i + 1, sigma)
      break 
    }
    #case where we need to pay out dividends to evaluate new stock price
    else{
      stock_price = new_q_price - dividend
    }
  }
  future_option_value = max(c(stock_price - E, 0))
  return(future_option_value - risk_free_rate(inflation_rate,period_to_maturity)*future_option_value)
}

#simulation looking at the case of 2 quarters to maturity and a stock price of 25 matching
option_vals = rep(NA, trials)
for (i in 1:trials){
  option_vals[i] = generate_option_val(stock_prices[1], r, periods_to_maturity[1],sqrt(sigma_squared))
}
print(paste("option value: ",mean(option_vals)))
print(paste("se of estimate: ",sd(option_vals)/(sqrt(trials))))
```

WORKS

```{r}
set.seed(400)
trials = 5000
periods_to_maturity = c(2,4,6,8,10,12,14,16,18,20)
stock_prices = c(25, 50, 75)
stockExerciseRatio = c(0.5, 1, 1.5)
E = 50
r = 0.015
sigma = 0.025 
dividend = 0.25

#chosen from October 2021, but others can be chosen to try to match historical data
yield_points = c(.09,.1,.11,.16,.36)
#yield_points = c(.04,.06,.9,.10,.14)
#inflation_rate = .1585
inflation_rate = 0.062

#rate to be used to evaluate the present day value of an investment
risk_free_rate = function(inflation_rate, periods_to_maturity){
  quarters = c(2,4,8,12,20)
  interpolate_yield_curve = approxfun(x = quarters, y = yield_points)
  return(inflation_rate - interpolate_yield_curve(periods_to_maturity))
}


# an individual evaluation of a trial for a single option valuation problem
generate_option_val = function(stock_price, r, periods_to_maturity, sigma){
  x_tilda = rnorm(periods_to_maturity, mean = 0, sd = 1)
  for (i in 1:periods_to_maturity){
    new_q_price = stock_price * exp((r - (sigma^2)/2) + sigma * x_tilda[i])
    #case where trial needs to be restarted from price at quarter t rather than quarter 1(see p. 329 of Boyle(1977))
    if (new_q_price -  dividend <= 0){
      stock_price = generate_stock_price(stock_price, r, periods_to_maturity - i + 1, sigma)
      break 
    }
    #case where we need to pay out dividends to evaluate new stock price
    else{
      stock_price = new_q_price - dividend
    }
  }
  future_option_value = max(c(stock_price - E, 0))
  return(future_option_value - risk_free_rate(inflation_rate,periods_to_maturity)*future_option_value)
}

#simulation looking at the case of 2 quarters to maturity and a stock price of 25 matching
option_vals = rep(NA, trials)
newmean = c()
newsd = c()
option_mean_25 = c()
option_sd_25 = c()

## need this to make variable option_vals again but empty so numbers dont add to each other
option_vals = rep(NA, trials)

for (l in 1:9){
  for (i in 1:trials){
    option_vals[i] = generate_option_val(stock_prices[1], r, periods_to_maturity[l],sqrt(sigma))
  }
  option_mean_25 = append(option_mean_25,mean(option_vals))
  newsd = sd(option_vals)/(sqrt(trials))
  option_sd_25 = append(option_sd_25,newsd)
}

## now this is for stock prices at 50

option_vals = rep(NA, trials)
newmean = c()
newsd = c()
option_mean_50 = c()
option_sd_50 = c()
for (l in 1:9){
  for (i in 1:trials){
    option_vals[i] = generate_option_val(stock_prices[2], r, periods_to_maturity[l],sqrt(sigma))
  }
  option_mean_50 = append(option_mean_50,mean(option_vals))
  newsd = sd(option_vals)/(sqrt(trials))
  option_sd_50 = append(option_sd_50,newsd)
}


## now this is for stock prices at 75
option_vals = rep(NA, trials)
newmean = c()
newsd = c()
option_mean_75 = c()
option_sd_75 = c()
for (l in 1:9){
  for (i in 1:trials){
    option_vals[i] = generate_option_val(stock_prices[3], r, periods_to_maturity[l],sqrt(sigma))
  }
  option_mean_75 = append(option_mean_75,mean(option_vals))
  newsd = sd(option_vals)/(sqrt(trials))
  option_sd_75 = append(option_sd_75,newsd)
}




periods_to_maturity =c(2,4,6,8,10,12,14,16,18)

df_25 <- data.frame(Periods_til_maturity = periods_to_maturity, Option_values=option_mean_25, Standard_deviation=option_sd_25, lower_95_CI = option_mean_25 - 2*sd(option_mean_25)/sqrt(trials), upper_95_CI = option_mean_25 + 2*sd(option_mean_25)/sqrt(trials))
df_25 = round(df_25, 4)
df_25

df_50 <- data.frame(Periods_til_maturity = periods_to_maturity, Option_values=option_mean_50, Standard_deviation=option_sd_50, lower_95_CI = option_mean_50 - 2*sd(option_mean_50)/sqrt(trials), upper_95_CI = option_mean_50 + 2*sd(option_mean_50)/sqrt(trials))
df_50 = round(df_50, 4)
df_50

df_75 <- data.frame(Periods_til_maturity = periods_to_maturity, Option_values=option_mean_75,Standard_deviation=option_sd_75, lower_95_CI = option_mean_75 - 2*sd(option_mean_75)/sqrt(trials), upper_95_CI = option_mean_75 + 2*sd(option_mean_75)/sqrt(trials))
df_75 = round(df_75, 4)
df_75
```

The simulation above was computed using the selected values of the underlying parameters:    
- S = 25, 50, 75 - current stock prices 
- E = 50    
- r = 0.015  (per period)    
- $sigma^2$ = 0.025  (per period)    
- D = 0.25  (per period)   
Monte Carlo estimates of options are calculated with even maturities 2 to 20       
Each estimate was obtained using 5000 trials where each trial corresponds to a possible stock path  


```{r}
ggplot(df_75, aes(x=Periods_til_maturity, y=Option_values)) + 
  geom_point()+
  xlab("Number of periods to maturity")+
  ylab("Option values by Crude Monte Carlo")+
  ggtitle("Scatter plot of values of options by period to maturity of at stock price of 75")


ggplot(df_50, aes(x=Periods_til_maturity, y=Option_values)) + 
  geom_point()+
  xlab("Number of periods to maturity")+
  ylab("Option values by Crude Monte Carlo")+
  ggtitle("Scatter plot of values of options by period to maturity of at stock price of 50")


ggplot(df_25, aes(x=Periods_til_maturity, y=Option_values)) + 
  geom_point()+
  xlab("Number of periods to maturity")+
  ylab("Option values by Crude Monte Carlo")+
  ggtitle("Scatter plot of values of options by period to maturity of at stock price of 25")

```


  
  
  
  
  
  
  
  
  
  

DOESnt WORK

```{r}
## copied code from above chunk to add to before making commit. Didnt want to mess up original code with a comit

trials = 5000
periods_to_maturity = c(2,4,6,8,10,12,14,16,18,20)


stock_prices = c(25, 50, 75)
stockExerciseRatio = c(0.5, 1, 1.25)
E = 50
r = 0.015
sigma = 0.025 
dividend = 0.25

#chosen from October 2021, but others can be chosen to try to match historical data
yield_points = c(.09,.1,.11,.16,.36)
inflation_rate = 0.062

#rate to be used to evaluate the present day value of an investment
risk_free_rate = function(inflation_rate, period_to_maturity){
  quarters = c(2,4,8,12,20)
  interpolate_yield_curve = approxfun(x = quarters, y = yield_points)
  return(inflation_rate - interpolate_yield_curve(period_to_maturity))
}
risk_free_rate

# an individual evaluation of a trial for a single option valuation problem
generate_option_val = function(stock_price, r, periods_to_maturity, sigma){
  x_tilda = rnorm(periods_to_maturity, mean = 0, sd = 1)
  for (i in 1:periods_to_maturity){
    new_q_price = stock_price * exp((r - (sigma^2)/2) + sigma * x_tilda[i])
    #case where trial needs to be restarted from price at quarter t rather than quarter 1(see p. 329 of Boyle(1977))
    if (new_q_price -  dividend <= 0){
      stock_price = generate_stock_price(stock_price, r, periods_to_maturity - i + 1, sigma)
      break 
    }
    #case where we need to pay out dividends to evaluate new stock price
    else{
      stock_price = new_q_price - dividend
    }
  }
  future_option_value = max(c(stock_price - E, 0))
  return(future_option_value - risk_free_rate(inflation_rate,periods_to_maturity)*future_option_value)
}

#simulation looking at the case of 2 quarters to maturity and a stock price of 25 matching
option_vals = rep(NA, trials)


## This is where code gets repetitive couldn't figure out a way to do a nested loop to efficiently make a table for evaluation of options

##Make empty variables to append values into a table then later clean up
newmean = c()
newsd = c()
option_mean_25 = c()
option_sd_25 = c()

## need this to make variable option_vals again but empty so numbers dont add to each other
option_vals = rep(NA, trials)
## cope for time period 2



for (l in 1:length(periods_to_maturity)){
  for (i in 1:trials){
  option_vals[i] = generate_option_val(stock_prices[1], r, periods_to_maturity[l],sqrt(sigma))


}
  option_mean = append(option_mean_25,mean(option_vals))
  newsd = sd(option_vals)/(sqrt(trials))
}

option_mean



for (i in 1:trials){
  option_vals[i] = generate_option_val(stock_prices[1], r, periods_to_maturity[1],sqrt(sigma_squared))

  newmean = mean(option_vals)
  newsd = sd(option_vals)/(sqrt(trials))
}

option_mean_25 = append(option_mean_25,newmean)
option_sd_25 = append(option_sd_25,newsd)

## copy of code this time for periods 4
option_vals = rep(NA, trials)
for (i in 1:trials){
  option_vals[i] = generate_option_val(stock_prices[1], r, periods_to_maturity[2],sqrt(sigma_squared))

  newmean = mean(option_vals)
  newsd = sd(option_vals)/(sqrt(trials))
}

option_mean_25 = append(option_mean_25,newmean)
option_sd_25 = append(option_sd_25,newsd)


## copy of code this time for periods 6
option_vals = rep(NA, trials)
for (i in 1:trials){
  option_vals[i] = generate_option_val(stock_prices[1], r, periods_to_maturity[3],sqrt(sigma_squared))

  newmean = mean(option_vals)
  newsd = sd(option_vals)/(sqrt(trials))
}

option_mean_25 = append(option_mean_25,newmean)
option_sd_25 = append(option_sd_25,newsd)


## copy of code this time for periods 8
option_vals = rep(NA, trials)
for (i in 1:trials){
  option_vals[i] = generate_option_val(stock_prices[1], r, periods_to_maturity[4],sqrt(sigma_squared))

  newmean = mean(option_vals)
  newsd = sd(option_vals)/(sqrt(trials))
}

option_mean_25 = append(option_mean_25,newmean)
option_sd_25 = append(option_sd_25,newsd)


## copy of code this time for periods 10
option_vals = rep(NA, trials)
for (i in 1:trials){
  option_vals[i] = generate_option_val(stock_prices[1], r, periods_to_maturity[5],sqrt(sigma_squared))

  newmean = mean(option_vals)
  newsd = sd(option_vals)/(sqrt(trials))
}

option_mean_25 = append(option_mean_25,newmean)
option_sd_25 = append(option_sd_25,newsd)


## copy of code this time for periods 12
option_vals = rep(NA, trials)
for (i in 1:trials){
  option_vals[i] = generate_option_val(stock_prices[1], r, periods_to_maturity[9],sqrt(sigma_squared))

  newmean = mean(option_vals)
  newsd = sd(option_vals)/(sqrt(trials))
}

option_mean_25 = append(option_mean_25,newmean)
option_sd_25 = append(option_sd_25,newsd)

```



